---
title: "Lab_6_Penguin_ANOVA"
author: "Eamon Devlin"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguin data
```{r, include= FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")
```

Our question

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update the species name using case_when

```{r}

counts <- with(penguins, table(species, sex))

counts

penguins2 <- penguins %>% 
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~"Chinstrap", 
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie")) %>% 
           select(-species) %>% 
           filter(sex == "FEMALE")
```

- Exploratory graphs
- Summary statistics
- Levene's tests for equal variances
- One way ANOVA
-